Optimizing Large Scale Imaging Surveys for a Retrospective Relative Photometric Calibration

### 2011-11-04 paper/self_calibration_I.tex
euclid: paper draft 0.8 complete
git-svn-id: svn+ssh://astrometry.net/svn/trunk/projects/euclid@19865 03a3cea6-2c03-0410-ac9b-9271b1c66c29

FROM HERE NEEDS TO BE UPDATED

The self-calibration procedure performs best with the Survey D dataset, followed by the Survey C dataset, then the Survey B dataset. The self-calibration procedure failed completely with the Survey A dataset and no solution was obtained.

With the Survey C and D datasets, the instrument response is accurately recovered. As can be seen from Figs. \ref{fig:C_result} and \ref{fig:D_result} (top) the fitted instrument response $f_\fit(\vec{x_j} | \vec{q_\fit})$ does not perfectly reproduce the true instrument response $f_\true(\vec{x} | \vec{q_\true})$, with residuals up to the $\pm 1~\%$ level. That said, through comparisons with the best fit possible with the eight order polynomial basis used in the self-calibration procedure $f_\basis(\vec{x_j} | \vec{q_\basis})$ (bottom), we find that majority of the remaining errors in our fits are due to limitations in the basis used. The residuals in this comparison are low; this can also be clearly seen in the factor of $\sim 10$ improvement in the Best-in-Basis Badness $B_\basis$ compared to the True Badness $B_\true$. The results from these two survey datasets are clear: if the correct redundancy is built into the survey, then self-calibration calibration can be used to accurately constrain the relative instrument response, with presumably the majority of the remaining errors coming from the limitations in the instrument response model fitted for.

The results from the self-calibration procedure with the Survey A and B datasets are also interesting.

tiling is crap

parts well connected work well

 when the fitted instrument response is compared to the best fit possible with the eight order polynomial basis used in the self-calibration procedure $f_\basis(\vec{x_j} | \vec{q_\basis})$ (bottom), we can see that the final solution is close

possible instrument response fit with the basis used $f_\basis(\vec{x_j} | \vec{q_\basis})$

The difference between the success of the self-calibration procedure with these four survey strategies is clear: with Surveys A no solution can be obtained, but with the Surveys B, C and D datasets the self-calibration procedure converges close to the correct solution. With the Survey C and D datasets the instrument response can be accurately fitted, close to the best fit possible with the basis used. The majority of the remaining errors in the fit come from the limitations of the basis used in the instrument response model. The calibration procedure with the Survey B dataset requires more iteration steps (4270) to converge to a final solution, and is less accurate than those found with the Survey C and D datasets, which require 28 and 10 iteration steps respectively.

The results from this investigation into simple survey strategies can be used to draw conclusions about the properties of surveys that make them good, or bad, for retrospective self-calibration. From these examples, I conclude that regular survey strategies with little overlap between adjacent pointings (Survey A) -- such as the Euclid Wide Survey -- do not allow for the resulting dataset to be self-calibrated effectively. It is therefore necessary to depart from the standard, regularly tiled observing strategies in the Euclid Deep Field. The two best survey strategies (C, D) both return the same sources to very different focal plane positions, and this is the key property which makes their datasets good for self-calibration. By doing so, many different focal plane positions are connected to each other with observations of the same source. This result must be considered when the Deep Field survey strategy is defined and as such I have applied calibration requirements that capture this on the strategy (see Subsection \ref{sec:deep_req}). In addition, I also conclude that if the correct redundancy is built into the survey strategy, the self-calibration procedure will be able to accurately constrain the relative instrument response, and therefore the Deep Field relative photometric calibration could be transfered effectively to the Wide Survey. The majority of the remaining errors will come from the limitations of the basis used to fit the instrument response model.

\section{Discussion}
In this paper, we have considered the optimization of survey strategies for calibrating the relative photometric response of an instrument, but we feel there are many more calibration parameters that can be constrained for this method. For example, the optical distortion of an instrument could also be constrained with such a method, although we concede that the properties of a survey strategy that makes them good for one calibration, may not be the same as that for another; ultimately a trade-off would need to be performed.

>>>>

The self-calibration procedure performs best with the Survey D dataset, followed by the Survey C dataset, then the Survey B dataset. The self-calibration procedure failed completely with the Survey A dataset and no solution was obtained. The number of iterations of the self-calibration procedure required to converge to the final solution also decreases as the quality of the final fit increases: Survey D required 9 iterations, Survey C 26, Survey B 4160 and the procedure was iterated XX times with Survey A without reaching a solution.

With the Survey C and D datasets, the instrument response is accurately recovered. As can be seen from Figures \ref{fig:C_result} and \ref{fig:D_result} (top) the fitted instrument response $f_\fit(\vec{x_j} | \vec{q_\fit})$ does not perfectly reproduce the true instrument response $f_\true(\vec{x} | \vec{q_\true})$, with residuals up to the $\pm 1~\%$ level. That said, through comparisons with the best fit possible with the eight order polynomial basis used in the self-calibration procedure $f_\basis(\vec{x_j} | \vec{q_\basis})$ (bottom), we find that majority of the remaining errors in our fits are due to limitations in the basis used. The residuals in this comparison are low; this can also be clearly seen in the factor of $\sim 10$ improvement in the Best-in-Basis Badness $B_\basis$ compared to the True Badness $B_\true$. The results from these two survey datasets are clear: if the correct redundancy is built into the survey, then self-calibration calibration can be used to accurately constrain the relative instrument response, with presumably the majority of the remaining errors coming from the limitations in the instrument response model fitted for.

The Survey A and B datasets also give interesting results. With Survey A, the self-calibration procedure did not tend to a solution. This, in itself, is an important result, as the naive tiling of the sky currently being considered for many large surveys will result in datasets that cannot be retrospectively self-calibrated. Minor changes to survey strategies, such as those in Survey C, can drastically improve the performance of the self-calibration procedure. Without these changes, vast amounts of calibration information is simply lost. We can also draw conclusions from the results obtained with Survey C: here the self-calibration procedure converges close to the correct solutions, albeit not as close as the Surveys C and B. Here, it is the quality of the fit over the focal plane that is interesting. From the residual plot against the best-in-basis instrument response $f_\basis(\vec{x_j} | \vec{q_\basis})$ (Fig \ref{fig:B_result} bottom right), we can see that the quality of the fit is rotationally symmetric, with the self-calibration procedure able to fit the instrument response well at the edge of the focal plane, but poorly at the center. This, along with the success of the procedure with the Surveys C and D, gives an insight into the key property of a survey that makes it good for self-calibration. Namely, the survey should return the same sources to many, very different focal plane positions. This allows the self-calibration procedure to link the different parts of the focal plane with each other through the observations of the same sources. Surveys C and D do this and the self-calibration procedure performs well with their dataset. Survey A does not do this; it only connects the outer edges of the focal plane to each other. There are no redundant measurements connecting the majority of the focal plane to other positions. This conclusion in confirmed with Strategy B. Here, positions at the outside of the focal plane are well connected to other positions and therefore the self-calibration procedure can well fit the instrument response in these regions. In the center of the focal plane, on which the same sources always fall, there are insufficient redundant measurements to allow for a good fit. Returning the same sources to very different focal plane positions is critical for the self-calibration to work effectively.

\section{Discussion}
From the simulations presented in this paper, we are draw two firm conclusions about the self-calibration of large-scale imaging surveys. The first is that, in agreement with previous work, we show that this technique is very powerful, allowing the relative instrument response of an imaging instrument to be accurately retrospectively constrained from the science images alone. The second is that there must be the correct redundancy built into the survey strategy. The same sources must be imaged at very different focal plane positions; regular tiling of the sky is useless for this kind of calibration.

In this paper, we have considered the optimization of survey strategies for calibrating the relative photometric response of an instrument, but we feel there are many more calibration parameters that can be constrained from this method. For example, the optical distortion of an instrument could also be constrained with such a method, although we concede that the properties of a survey strategy that makes them good for one calibration, may not be the same as that for another; ultimately a trade-off would need to be performed.
